\begin{thebibliography}{52}

\bibitem{a-b-b:CBA-m-w-b} 
Alexander, S.; Berg, D.; Bishop, R., 
\textit{Geometric curvature bounds in Riemannian manifolds with boundary}. 
Trans. Amer. Math. Soc., 339 (1993), 703--716.

\bibitem{a-b:h-c}  
Alexander, S.; 
Bishop, R., 
\textit{The Hadamard--Cartan theorem in locally convex spaces}. 
Enseignement Math., 
36 
(1990), 
309--320.

\bibitem{a-b:warped-} 
Alexander, S.; 
Bishop, R., 
\textit{Warped products of Hadamard spaces}.
Manuscripta Math. 96 (1998), no. 4, 487--505.


\bibitem{a-b:warped} 
Alexander, S.; 
Bishop, R., 
\textit{Curvature bounds for warped products of metric spaces}.
GAFA,
14 (2004),
\#6, 
1143--1181.

\bibitem{a-b:warped+} 
Alexander, S.; 
Bishop, R., 
\textit{Warped products admitting a curvature bound}.
 Adv. Math. 303 (2016), 88--122.

\bibitem{AKP}
Alexander, S.;
Kapovitch, V.;
Petrunin, A.,
\textit{Alexandrov geometry}.
 \texttt{http://www.math.psu.edu/petrunin/}
 
\bibitem{alexandrov:def}
\begin{otherlanguage}{russian}
 Александров, А. Д.,
\textit{Внутренняя геометрия произвольной выпуклой поверхности}.
Доклады Академии наук СССР, 32 (1941) № 7, 467---470.
\end{otherlanguage}
%Alexandroff, A.
%The inner geometry of an arbitrary convex surface.
%C. R. (Doklady) Acad. Sci. URSS (N.S.) 32, (1941), 467--470.

\bibitem{alexandrov:strong-angle} 
\begin{otherlanguage}{russian}
Александров, А. Д.,
\textit{Одна теорема о треугольниках в метрическом пространстве и некоторые ее приложения}.
Труды МИАН СССР, 38 (1951), 5---23;
\end{otherlanguage}
translated into German and combined with more material in \cite{alexandrov:devel}.


\bibitem{A-ruled}
\begin{otherlanguage}{russian}
Александров, А. Д., 
\textit{Линейчатые поверхности в метрических пространствах}.
Вестник ЛГУ 2 (1957): 15---44.
\end{otherlanguage}
%Alexandrov, A. D. ``Ruled  surfaces  in  metric  spaces,'' Vestnik Leningrad. Univ., 12:5-26, 1957 (Russian).

\bibitem{alexandrov:devel} Alexandrow, A. D.,  \textit{\"Uber eine Verallgemeinerung der Riemannschen Geometrie}.
Schriftenreihe Inst. Math. 
1 
(1957), 
p. 33--84.

\bibitem{ADG} Ancel, F.; Davis, M.; Guilbault, C.,
\textit{CAT(0) reflection manifolds} .
Geometric topology, 
441--445.

\bibitem{ballmann:cartan-hadamard}
Ballmann, W., 
\textit{Singular spaces of non-positive curvature}. 
Sur les Groupes Hyperboliques d'apr\'{e}s Mikhael Gromov, Chapter 10, Progr. Math. 83, Birkh\"auser, Boston, 1990. 

\bibitem{ballmann:lectures}
Ballmann, W., \textit{Lectures on Spaces of Nonpositive Curvature}.
with an appendix by Misha Brin, 1995. 

\bibitem{BHV} 
Billera, L.; 
Holmes, S.; 
Vogtmann, K., 
\textit{Geometry of the space of phylogenetic trees}.
Adv. in Appl. Math. 
27 
(2001), 
no. 4, 
733--767.

\bibitem{bishop:jordan}
Bishop, R., 
\textit{The intrinsic geometry of a Jordan domain}. 
Int. Electron. J. Geom. 
1 
(2008), 
no. 2, 
33--39.

\bibitem{blaschke} 
Blaschke, W., 
\textit{Kreis und Kugel}. 1916.

\bibitem{bowditch} 
Bowditch, B. H.,
\textit{Notes on locally CAT(1) spaces}.
in ``Geometric group theory'', 
de Gruyter (1995) 1--48.

\bibitem{bowditch-2020} Bowditch, B. H.)
\textit{Median and injective metric spaces.}
Math. Proc. Cambridge Philos. Soc. 168 (2020), no. 1, 43--55.

\bibitem{BH}
Bridson, M., 
Haefliger, A.,
\textit{Metric Spaces of Non-positive Curvature}. 1999.

\bibitem{burago-icm} 
Burago, D., 
\textit{Hard balls gas and Alexandrov spaces of curvature bounded above}.
Proceedings of the International Congress of Mathematicians, Vol. II, 
Doc. Math. 1998, 
Extra Vol. II, 
289--298.

\bibitem{BBI} 
Burago, D.; 
Burago, Yu.; 
Ivanov, S., 
\textit{A course in metric geometry}. 2001. 

\bibitem{BFK}
Burago, D.; 
Ferleger, S.; 
Kononenko, A.,
\textit{Uniform estimates on the number of collisions in semi-dispersing billiards}.
Annals of Math.(2) 147 (1998), 695--708.

\bibitem{BFK-2} 
Burago, D.; 
Ferleger, S.;
Kononenko, A., 
\textit{Topological entropy of semi-dispersing billiards}.
Ergodic Theory and Dynamical Systems 
18.04 
(1998), 
791--805.

\bibitem{BGS}
Burago, D.; Grigoriev, D.; Slissenko, A.
\textit{Approximating shortest path for the skew lines problem in time doubly logarithmic in 1/epsilon}. 
Theoret. Comput. Sci. 
315 (2004), 
no. 2-3, 
371--404. 

\bibitem{burago-ivanov}
Burago, D., Ivanov, S. 
\textit{Examples of exponentially many collisions in a hard ball system}.
\texttt{arXiv:1809.02800 [math.DS]}.
 

\bibitem{busemann-CBA}
Busemann, H., 
\textit{Spaces with non-positive curvature}.
Acta Math. 
80, 
(1948) 
259--310.

\bibitem{buyalo}
Buyalo, S. V.,
\textit{Volume and fundamental group of a manifold of nonpositive curvature}.
Mathematics of the USSR-Sbornik 50, no. 1 (1985): 137--150.

\bibitem{cartan}
Cartan, \'E.,
\textit{Le\c{c}ons sur la G\'eom\'etrie des Espaces de Riemann}. 1928.

\bibitem{charney-davis-93}
Charney, R.;
Davis, M.,
\textit{Singular Metrics of Nonpositive Curvature on Branched Covers of Riemannian Manifolds}.
American Journal of Mathematics, 
Vol. 115, 
No. 5 
(Oct., 1993), 
929--1009.

\bibitem{charney-davis}
Charney, R.; Davis, M., 
\textit{Strict hyperbolization}.
Topology 
34.2 
(1995) 
329--350.

\bibitem{daverman}
Daverman, R. J.
\textit{Decompositions of manifolds}. 1986.

\bibitem{davis-noneuclidean}
Davis, M., 
\textit{Groups generated by reflections and aspherical manifolds not covered by Euclidean space}.
Ann. of Math. 
(2) 117 
(1983), 
no. 2, 
293--324. 

\bibitem{davis-survey} 
Davis, M.,
\textit{Exotic aspherical manifolds}.
Topology of high-dimensional manifolds, 
No. 1, 2 (Trieste, 2001), 371--404, ICTP Lect. Notes, 9, 2002.



\bibitem{davis-januszkiewicz} 
Davis, M.; 
Januszkiewicz, T., 
\textit{Hyperbolization of polyhedra}. 
J. Differential Geom. 
34 
(1991), 
no. 2, 
347--388.

\bibitem{davis-januszkiewicz-lafont} 
Davis, M.; 
Januszkiewicz, T.; 
Lafont, J.-F.,
\textit{4-dimensional locally CAT(0)-manifolds with no Riemannian smoothings}. 
Duke Math. J. 
161 
(2012), 
no. 1, 
1--28.

\bibitem{edwards} 
Edwards, R.,
\textit{The topology of manifolds and cell-like maps}.
Proceedings of the International Congress of Mathematicians (Helsinki, 1978), pp. 111--127, 1980.

\bibitem{frolik} 
Frol\'{\i}k, Z., 
\textit{Concerning topological convergence of sets}.
Czechoskovak Math. J. 
10 
(1960), 
168--180.

\bibitem{galperin} 
\begin{otherlanguage}{russian} 
Гальперин, Г.~А.
\textit{О системах локально взаимодействующих и отталкивающихся частиц, движущихся в~пространстве}
Тр. ММО
43
(1981)
142--196
\end{otherlanguage}

\bibitem {gromov:hyp-groups}
Gromov. M.,
\textit{Hyperbolic groups}. 
Essays in Group Theory,
Math. Sciences Research Institute Publications 8,
1987, 75--264.

\bibitem{gromov:SaGMC}
Gromov, M.,
\textit{Sign and geometric meaning of curvature}.
1994.

\bibitem{hadamard} 
Hadamard, J.,
\textit{Sur la forme des lignes g\'eod\'esiques \'a l'infini et sur les g\'eod\'esiques des surfaces r\'egl\'ees du second ordre}.
Bull. Soc. Math. France 
26 
1898), 
195--216.

\bibitem{hatcher}
Hatcher, A., 
\textit{Notes on Basic 3-Manifold Topology}.
a draft available at \verb+http://www.math.cornell.edu/~hatcher/+.

\bibitem{hass} 
Hass, J.,
\textit{Bounded 3-manifolds admit negatively curved metrics with concave boundary}.
J. Differential Geom. 
40 (1994), 
no. 3, 
449--459. 

\bibitem{hausdorff}
Hausdorff, F., 
\textit{Grundz\"uge der Mengenlehre}. 
1914.

\bibitem{kervaire} 
Kervaire, M., 
\textit{Smooth homology spheres and their fundamental groups}.
Trans. Amer. Math. Soc. 
144 
(1969),
67--72.

\bibitem{menger} 
Menger, K.
\textit{The formative years of Abraham Wald and his work in geometry}.
Ann. Math. Statistics 23 (1952), 14--20.

\bibitem{lebedeva-petrunin-poly} 
Lebedeva, N.; 
Petrunin, A., 
\textit{Local characterization of polyhedral spaces}.
Geom. Dedicata 179 (2015), 161--168.

\bibitem{lytchak-wenger}
Lytchak, A.; Wenger, S.,
\textit{Isoperimetric characterization of upper curvature bounds}.
Acta Math. 221 (2018), no. 1, 159--202.

\bibitem{lytchak-stadler}
Lytchak, A.; Stadler, S. 
\textit{Conformal deformations of CAT(0) spaces}
Math. Ann. (2018). 

\bibitem{mazur}
Mazur, B.,
\textit{A note on some contractible 4-manifolds}.
Ann. of Math. (2) 
73 
(1961),
221--228.

\bibitem{mangoldt}
von Mangoldt, H., 
\textit{Ueber diejenigen Punkte auf positiv gekr\"ummten Fl\"achen, welche die Eigenschaft haben, dass die von ihnen ausgehenden geod\"atischen Linien nie aufh\"oren, k\"urzeste Linien zu sein}.
J. Reine Angew. Math. 91 (1881), 23--53.

\bibitem{nazarov}
Nazarov, F.,
\textit{Intrinsic metric with no geodesics}.
MathOverflow \texttt{http://mathoverflow.net/q/15720}.

\bibitem{panov-petrunin:sweeping}
Panov, D.; Petrunin, A.,
\textit{Sweeping out sectional curvature}. 
Geom. Topol. 
18 
(2014), 
no. 2, 
617--631. 

\bibitem{panov-petrunin:ramification}
Panov, D.; Petrunin, A.,
\textit{Ramification conjecture and Hirzebruch's property of line arrangements}.
to appear in Compositio Mathematica.

\bibitem{petrunin-metricmin} 
Petrunin, A.,
\textit{Metric minimizing surfaces}.
Electron. Res. Announc. Amer. Math. Soc. 
5 
(1999), 
47--54.

\bibitem{petrunin-quest}  
Petrunin, A. 
\textit{A quest for 5-point condition a la Alexandrov}.
St. Petersburg Math. J. 29 (2018), no. 1, 223--225.

\bibitem{petrunin-stadler-saddle}  
Petrunin, A.; 
Stadler, S., 
\textit{Monotonicity of saddle maps}.
\texttt{arXiv:1707.02367 [math.DG]}, to appear in Geom. Dedicata.

\bibitem{petrunin-stadler}  
Petrunin, A.; 
Stadler, S., 
\textit{Metric minimizing surfaces revisited}.
\texttt{arXiv:1707.09635 [math.DG]}.

\bibitem{reshetnyak:glue}
Reshetnyak, Yu. G.,
\textit{On the theory of spaces of curvature not greater than $K$}.
Inextensible mappings in a space of curvature
no greater than $K$, Siberian Math. Jour. 9 (1968), 683--689.

\bibitem{reshetnyak:survey}
%\begin{otherlanguage}{russian}
%Решетняк, Ю. Г.
%\textit{Двумерные многообразия ограниченной кривизны}
%Геометрия~--~4 (1989), 7--189
%\end{otherlanguage}
Reshetnyak, Yu. G., 
\textit{Two-dimensional manifolds of bounded curvature}.
Geometry, IV, 3--163, 245--250, Encyclopaedia Math. Sci., 70, Springer, Berlin, 1993.

\bibitem{rolfsen} 
Rolfsen, D.,
\textit{Strongly convex metrics in cells}.
Bull. Amer. Math. Soc. 74 (1968) 171–175.

\bibitem{rinow}
Rinow, W., 
\textit{Die innere Geometrie der metrischen Raume}.
1961.

\bibitem{shefel-2D} 
%\begin{otherlanguage}{russian}
%Шефель, С. З.,
%\textit{О седловых поверхностях ограниченной спрямляемой кривой.}
%Доклады АН СССР, 162 (1965) №2, 
%294---296.
%\end{otherlanguage}
%English translation: 
\v{S}efel', S.
\textit{On saddle surfaces bounded by a rectifiable curve}.
Soviet Math. Dokl. 6 (1965), 684--687.


\bibitem{shefel-3D} 
\begin{otherlanguage}{russian}
Шефель, С. З., 
\textit{О внутренней геометрии седловых поверхностей}.
Сибирский математический журнал, 5 (1964), 1382---1396.
\end{otherlanguage}
%\v{S}efel', S., 
%\textit{On the intrinsic geometry of saddle surfaces,} Sibirsk. Mat. \v{Z}. 
%5 
%(1964), 
%1382--1396

\bibitem{stadler}
Stadler, S.,
\textit{An obstruction to the smoothability of singular nonpositively curved metrics on 4-manifolds by patterns of incompressible tori}.
GAFA
25 (2015), 
no 5, 
1575--1587.

\bibitem{stone}
Stone, D.,
\textit{Geodesics in piecewise linear manifolds}.
Trans. Amer. Math. Soc. 
215 
(1976), 
1--44.

\bibitem{thurston} Thurston, P.,
\textit{CAT(0) 4-manifolds possessing a single tame point are Euclidean}. 
J. Geom. Anal. 
6 (1996), 
no. 3, 
475--494 (1997). 

\bibitem{tietze} Tietze, H.,
\emph{\"Uber Konvexheit im kleinen und im gro\ss{}en und \"uber gewisse den Punkten einer Menge zugeordnete Dimensionszahlen}.
Math. Z. 28 (1928), no. 1, 697--707. 

\bibitem{wald} Wald, A., \textit{Begr\"undung eiiner Koordinatenlosen Differentialgeometrie der Fl\"achen}.
Ergebnisse eines mathematischen Kolloquium
6  (1935), 24--46.

\bibitem{wijsman} Wijsman, R. A., 
\textit{Convergence of sequences of convex sets, cones and functions. II}. 
Trans. Amer. Math. Soc. 
123 (1966) 32--45.


\end{thebibliography}
